---
title: "House of Representatives - Preferences"
resource_files:
  - vignetttes/house-preference-flow*.png
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{House of Representatives - Preferences}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/house-preference-flow",
  message = FALSE, warning = FALSE, error = FALSE
)
library(ggplot2)
library(auspol)

```

Preferences are a key aspect of Australia's preferential voting system.  **auspol** includes two functions to interact with the preference distribution data:

- get_house_preferences()
- house_preference_flow_data()
- house_preference_flow_plot()
- get_house_2PF()
- get_house_2PP()

## What is this?

If you are unfamiliar with the Australian electoral system and preferential voting, please look at this [explainer](https://www.aec.gov.au/learn/preferential-voting.html) before proceeding.

## Getting the data

get_house_preferences() - provides an interface to access preference flows for a particular division (electorate) on a given year, as published on the AEC's website. For instance:

```{r}
get_house_preferences("Wills",2019)

```

Results can be also retrieved disaggregated by polling place:

```{r}
get_house_preferences("Wills",2019,aggregation = FALSE) |> head(10)
                      
```

It is also possible to filter by Polling Place Name (PPNm in the source):

```{r}
get_house_preferences("Wills",2019, polling_places=c("ABSENT"),aggregation = FALSE) |> head(10)

```

Built on top of the basic functions,  preference_flow_data() - provides an interface to access preference flows for a particular division (electorate) in a given year. This function returns a list, in which each element is a data frame with the results for each counting round. In its most basic use:

```{r}
house_preference_flow_data(division = "Wills", year=2019)
```

The function also includes other variables, which are better explained through its use as part of the next function,  house_preference_flow_plot().

## Plotting Preference flows.

Perhaps it is most interesting to see how each round of counting assigns votes until a candidate is chosen. This package comes with the convenience function house_preference_flow_plot(), which calls preference_flow_data() and plots the results in an alluvial format. Using the previous example:

```{r plot1}
house_preference_flow_plot(division = "Wills",year=2019) +
  labs(title="Preference Flow for Wills, 2019")
```

By default the vote is present as a percentage, which can be changed by changing the value of *var*:

```{r plot2}
house_preference_flow_plot(division = "Wills",year=2019,var="Preference Count") +
  labs(title="Preference Flow for Wills, 2019")

```

For analysis, it may be interesting to remove some parties from the initial round - e.g. to see how votes flow to the finalists. This can be done using *exclude_parties*:

```{r plot3}
house_preference_flow_plot(division = "Warringah",year=2022) +
  labs(title="Preference Flow for Warringah, 2022")

```

```{r plot4}
house_preference_flow_plot(division = "Warringah",year=2022, exclude_parties = c("LP","IND-STEGGALL")) +
  labs(title="Flow of preference votes to Liberal and Teal in Warringah, 2022")

```

Like the other plotting functions, it is possible to override the colour scheme by providing an additional named vector:

```{r plot5}
#adding teal
house_preference_flow_plot(division = "Warringah",year=2022, 
                     exclude_parties = c("LP","IND-STEGGALL"),
                     extra_colours = c("IND-STEGGALL"="#008080")) +
  labs(title="Flow of preference votes to Liberal and Teal in Warringah, 2022")
```

Similarly, it is possible to get the results of house_preference_flow_plot() directly, via *include_data*:

```{r}
p <- house_preference_flow_plot(division = "Warringah",year=2022, 
                     exclude_parties = c("LP","IND-STEGGALL"),
                     extra_colours = c("IND-STEGGALL"="#008080"),
                     include_data=TRUE
                     )

p$source_data[[1]]
```

# Flow from primary vote to last round

Although the last plot can be informative, The AEC also records how o understand how the votes have flowed from primary votes to the two last candidates. This can be accessed through *get_house_2PF()*

```{r}
get_house_2PF(division = "Lingiari",year=2013,aggregation = TRUE)
```

These results can be charted using *house_2PF_plot()*. This function can present the results either in an alluvial or bar chart format.

```{r plot6}
house_2PF_plot("Lingiari",2013,plot_format = "alluvial")
```

```{r plot7}
house_2PF_plot(division="Burt",
               year=2022,
               plot_format = "bar")

```

# Two-party preferred

The AEC also publishes  [Two-Party-Preferred](https://en.wikipedia.org/wiki/Two-party-preferred_vote) summaries ([Coalition](https://en.wikipedia.org/wiki/Coalition_(Australia)) against Labor). These are available through * get_house_2PP()*, which allows filtering by division and year, and can be displayed by division or polling place

```{r}
get_house_2PP(division = "Fraser",year=2004,aggregation = TRUE)
```

The package includes two plots to visualise this data. The first (*house_2PP_comparison_plot()*),  displays the results for a given election, for one or many electorates.

```{r plot8}
house_2PP_comparison_plot(year=2022,state="VIC")
```

The second function - *house_2PP_historical_plot()* - shows the historical results for a given division.

```{r plot9}
house_2PP_historical_plot(division="Aston")
```
