---
title: "House of Representatives - Primary Vote"
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{House of Representatives - Primary Vote}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "man/figures/house-primary-vote",
  message = FALSE, warning = FALSE, error = FALSE
)

library(ggplot2)
library(auspol)
get_house_primary_vote()

```

**auspol** includes two functions to interact with the preference distribution data:

- get_house_primary_vote()
- house_primary_vote_summary()
- house_primary_comparison_plot()
- house_primary_historic_plot()

## What is this?

If you are unfamiliar with the Australian electoral system and preferential voting, please look at this [explainer(https://www.aec.gov.au/learn/preferential-voting.html) before proceeding.


## Getting the data

*get_house_primary_vote()* is the basic function to retrieve primary vote data as published by the AEC. Without any arguments, it will delivery all the results for all elections, but it comes with parameters to facilitate filtering.
For instance, to get the results for Brisbane for 2022:

```{r}
get_house_primary_vote(division="Brisbane",year=2022)
```

Both parameters can include more than one value, e.g.

```{r}
get_house_primary_vote(division="Brisbane",year=c(2019,2022))
```

```{r}
get_house_primary_vote(division=c("Brisbane","Perth"),year=c(2019,2022))
```

By default, the results are presented by polling place, with it is possible to aggregate them.

```{r}
get_house_primary_vote(division=c("Brisbane","Perth"),year=c(2019,2022),aggregation = TRUE)
```


```{r}
get_house_primary_vote(division=c("Brisbane","Perth"),year=c(2019,2022), polling_places = c("Yokine North"))
```

It is also possible to restrict the results to selected polling places



Additionally, it is possible to select one or more states instead of a group divisions, e.g.:

```{r}
get_house_primary_vote(state=c("TAS"),year=c(2019,2022),aggregation = TRUE)
```

It is also possible to filter results by one or more parties:

```{r}
get_house_primary_vote(state=c("NT"),year=c(2019,2022),aggregation = TRUE, party_abb=c("ALP","CLP"))

```


*house_primary_vote_summary()* builds on the basic function and summarise data
.
```{r}
house_primary_vote_summary(division = "Brisbane", year=2022)
```

Using the previous filters,  it is possible to get ad-hoc summaries, for instance - all the ALP vote in Queensland in 2022, or the historic Liberal vote in Franklin.

```{r}

house_primary_vote_summary(state = "QLD", year=2022,parties="ALP")

```



```{r}
house_primary_vote_summary(division="Franklin",parties="LP")
```


## Plotting 

### Historic Trends


The first plotting convenience function in this packages allows to compare the evolution of primary voting across time. This function relies on house_primary_summary and uses many of its options. It's first use is to represent party trends in one electorate:

```{r}
house_primary_historic_plot("Canberra")
```

As in many cases there are many minor parties, it is sometime useful just to focus in a number of parties. This function allow to filter by a number of parties or by filtering by the most voted in a certain year. In both cases it is possible to consolidate the others votes.



```{r}
house_primary_historic_plot("Canberra", parties =3, parties_year = 2022, include_others = TRUE )

```

Finally, it is possible to aggregate party acronyms - sometimes a same party has changed named or registered differently

```{r}

house_primary_historic_plot(division="Brisbane",parties=5,
                            merge_parties = list(LNP=c("LNP","LNQ","LP"))) 

```


