---
title: "House of Representatives - Get the data"
resource_files:
  - vignetttes/house-preference-flow*.png
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{House of Representatives - Get the data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/house-get-data-",
  message = FALSE, warning = FALSE, error = FALSE
)
library(ggplot2)
library(auspol)

```


Although this package contains many convenience functions to process and display the election data, there are probably many cases where the data in its original format is preferable - especially by and advanced analysis doing ad-hoc analysis. For these reason, the functions that read the data in its "original" form are also exported for use. These are:

- get_house_primary_vote()
- get_house_preferences()
- get_house_2PF()
- get_house_2PP()
- get_house_MPs()
- get_house_turnout()


Please note that the presented here is shown as downloaded from the AEC's website.

## Primary Vote

It is possible to get all primary vote results for all electorates in all electon from 2044, by running *get_house_primary_vote()*. More usefully, the funcion allows to filter by a number of electorates for a numbers of years, as shown in the below example:

```{r}
get_house_primary_vote(division = c("Cowan","Curtin"),year=c(2019,2022))
```

The function allows also to narrow down to specific electorates. For instance, to check the primary vote for ALP candidates, for both electorates in polling places named "Perth":


```{r}
library(dplyr)

get_house_primary_vote(division = c("Cowan","Curtin"),
                       year=c(2019,2022),
                       polling_places = "Perth") |>
  filter(PartyAb=="ALP")
```

The function also can present the division totals, for example:

```{r}
get_house_primary_vote(division = c("Cowan","Curtin"),
                       year=c(2019,2022), 
                       aggregation = TRUE)
```

## Preferences

To retrieve preferences for one division for a particular year, the function is called *get_house_preferences()*

```{r}
get_house_preferences("Wills",2019) |> head(10)

```


Similarly to the primary vote function, results can be aggregated and filtered by polling place:

```{r}
get_house_preferences("Wills",2019,
                      aggregation = TRUE)

```


```{r}
get_house_preferences("Wills",2019, polling_places=c("ABSENT")) |> 
  head(10)
```
The AEC also provides a list of how the primary votes have flowed to the two finalists. The can be retrieved with *get_house_2PF*. Being related to the previous function, it also has the same parameters.

```{r}
get_house_2PF(division="Jagajaga",year=2013,aggregation = TRUE)
```
The last preference function is *get_house_2PP()*, which provides the [Two-Party-Preferred](https://en.wikipedia.org/wiki/Two-party-preferred_vote) summaries as recorded by the AEC (usually, Coalition vs. ALP.)

```{r}
get_house_2PP(division = "Indi",
              year=2016,
              aggregation = TRUE)
```



## Elected MPs

The function *get_house_MPs()* retrieves the list of elected candidates. Without parameters it returns the complete lists (all divisions, for all elections), but similarly to the previous functions it allows for results to be filtered by divisions and years

```{r}
get_house_MPs(division = c("Melbourne","Cooper"),
              year = c(2019,2022))
```

## Turnout

The function *get_house_turnout()* retrieves turnout data as published by the AEC. It can be filtered by division and year

```{r}
get_house_turnout(year=2019)
```





