---
title: "Elected MPs"
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Elected MPs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/elected-",
  message = FALSE, warning = FALSE, error = FALSE
)

library(ggplot2)
library(auspol)
get_house_primary_vote()

```

This package contains the following functions to retrieve and visualise the election results in the House of Representatives.

- get_house_MPs()



## Getting the data

*get_house_MPs()*  retrieves the list of elected MPs. When used without any parameters, it will retrieve all elected MPs for all divisions, in all the elections in the package. However,data can be filtered by division and election year.


```{r}
get_house_MPs(division="Bennelong")
```


```{r}
get_house_MPs(year=2013)
```
## Plotting

This package also contains  functions to represent election results graphically. The first function presents and overall count of votes per party on a given year. For example, for 2013.

```{r plot1}
house_results_tally(2013)
```

Parties can be group together - i.e. putting all Coalition votes together


```{r plot2}
house_results_tally(2013, merge_parties = list(COAL=c("CLP","LP","LNP","NP")))

```

Furthermore, independents can be individualised - for instance for the 2022 elections:

```{r plot3}
house_results_tally(2022, 
                    individualise_IND = TRUE,
                    merge_parties = list(COAL=c("CLP","LP","LNP","NP"),
                                         TEAL=c("IND-DANIEL","IND-RYAN",
                                                "IND-SCAMPS","IND-CHANEY",
                                                "IND-SPENDER","IND-STEGGALL",
                                                "IND-HAINES","IND-TINK")
                                         ))

```

Finally, there is also a function to plot the historical seats per party


```{r plot4}
house_results_historic(merge_parties = list(COAL=c("CLP","LP","LNP","NP")),
                       parties =c("COAL","ALP","GRN"),
                       include_other=TRUE)
```

