---
title: "Listing the essential data"
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Listing the essential data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/lists",
  message = FALSE, warning = FALSE, error = FALSE
)

library(ggplot2)
library(auspol)
get_house_primary_vote()

```

To get started, it is good to know in advance the division to analyse, parties in the race, etc. The *list_* family of functions will help you to get to bearings before getting and analysing the data. The functions include here are:

- list_years()
- list_divisions()
- list_parties()
- list_polling_places()

## Years and divisions

To figure out the election results included in the package, just run:

```{r}
list_years()

```

To retrieve the list of division for all the years, you can run *list_divisions*:

```{r}
list_divisions()
```


This function allows for some filtering by using a list containing for any of the columns included:

```{r}
list_divisions(filter=list(DivisionNm=c("Batman","Cooper")))
```

## Political Parties

It is recommend to explore the list of parties that have participated in the elections in the data. For this **

```{r}

list_parties()

```


Due to changes in parties and the federal nature of the country, parties may have different names in different places and they may change over time. In addition to a filter argument (like in list_divisions), this function also allows to filter by party names using  a [regular expression](https://en.wikipedia.org/wiki/Regular_expression), e.g.:

```{r}
list_parties(party_regex = "Greens")
```


## Polling places

Finally, it is possible to list all polling places, showing their active years. For exmple, for all the electorates in Bennelong

```{r}
list_polling_places(filters=list(DivisionNm="Bennelong"))
```




